\documentclass[twoside,11pt,a5paper]{memoir}

\usepackage[top=1.1in,bottom=0.9in]{geometry}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{pdfpages}

% Preamble
\usepackage[osf]{libertine}
\usepackage{microtype}

\nouppercaseheads


% Information
\title{My Life in and around\\Thorn Moor and Beyond}
\author{Mike Saffin}
\date{}


% DOCUMENT
\begin{document}

% Front cover
% \includepdf{pictures/front-cover.pdf}

\maketitle

\chapter*{Old Devon verse}
\input{content/14-verse}

\chapter*{Foreword}

I apologise to anyone who, having read this book, has walked away thinking,
``this man is arrogant and talks about himself''. It is not intended to read
that way, as in the past 6 years I have been asked to put my life on paper
before it is lost. So I put pen to paper in the full knowledge that I have no
idea how to write a book or to put any events into the correct grammar
context! I just set forth! What is correct grammar? Define grammar --- grammar
teaches us the proper arrangement of words according to the ideals and dialects
of any particular kingdom of people. It teaches us to speak and write with
precision. Agreeable to question authority, I am a Devonian, born and bred in
Devon, so the West Country culture sits squarely on my shoulders.

I thank my wife, Vera, for standing and supporting me throughout my many years
of projects. Without that support there would be nothing to write about, and I
also give big thanks to my lifelong friend Patience, who took away my scribbles
to be formed into this booklet.

\bigskip

\emph{Mike Saffin, October 2022}

\chapter{}
\input{content/1}
\input{content/2}
\chapter{}
\input{content/3}
\chapter{}
\input{content/4}
\chapter{}
\input{content/5}
\chapter{}
\input{content/6}
\input{content/7}
\chapter{}
\input{content/8}
\input{content/9}
\chapter{}
\input{content/10}

% Pictures
\clearpage
\input{content/pictures}

\chapter{}
\input{content/11}
\chapter{}
\input{content/12}
\chapter{}
\input{content/13}

% Back cover
% \includepdf{pictures/back-cover.pdf}

\end{document}
