%!TEX root = ../thorn-moor.tex

My Uncle Ronald once said, ``The trouble with folks moving down to Devon is that
they think we Devonians are daft because we have no sense!''. How true, as when
I was a young man I set out with one aim in view to finish the project only to
find that a journey down the other road would have achieved a better result ---
a fault that holds fast in future years. I decided to build a double-seated
canoe using plans from a practical woodworking magazine. The length of the
designed vessel was 17' 6'' to be built in my garage in a very cold winter
season --- this was mistake number one. I cut the length from 17' 6'' to 17' to
accommodate the project in the garage enabling the door to close and lock ---
mistake number two. To upgrade the marine ply from 3mm to 4mm, the hull spares
were increased in size. On completion the vessel looked solid but when launched
it immediately gave the impression that things were heavier than required and
the removal of the 6'' waterline length meant the hull sat lower in the water.
Nevertheless my friend and I entered into the Exeter Canoe Club race. Despite
being left at the post to start with, after half an hour of paddling we began
to catch the competitors and overtook at regular intervals being very concerned
that it was like pulling a log through the water rather than a responsive
vessel. We reached Double Locks, turned and kept up our relentless paddling,
finally reaching the finish at Exeter Basin. We were congratulated on achieving
2nd position in the touring canoe section. Congratulations were quickly
followed by reality as we wondered if there were only two touring canoes! A
further follow up from there is to never look on one side only. Turn the coin
over and look at the tails as well as the heads --- that applies to every
avenue of problems presented.

I have always been active in my garage and, as life has taken a turn, have been
given a 1937 metal lathe with multiple-speed pulley drives and a belt. I have
dismantled this system and have converted a three-phase motor to interphase
with a multi-speed computer system, thus eliminating the multiple pulley
system. I turned away from engines and gears and made staves and wooden and
brass fittings for banners, all used for parades in Plymouth. I have taught
myself leather work and made sword belts and frogs for parades. I have often
wondered, when sewing up a thatcher's hook pouch using the saddle stitch, if I
had learned this trade and followed the occupation and become a leather worker,
would I be content in my peaceful workshop, or would I yearn for something more
active as I did at Thorn Moor in my early days?

Going back to the police hilltop radio stations I mentioned previously, I
started one day to collect an architect from Hilltop in Plymouth and visited a
generator unit that is powered in a Devon and Cornwall Police hilltop radio
station. My companion was named Tom Jane. We knew each other well as we had
carried out many of such inspections, so conversation was always great, but a
happening occurred which affected both of us after we looked at the generator
and started our quite long descent to return. I was driving quite slowly so as
not to cause damage, as the path was very deeply rutted with small loose
boulders. Looking in the distance I could see a spiralling cloud of dust moving
very fast. Tom's voice said in alarm, ``Better pull in here where there is a
wide area''. We stopped and waited. The drive was in another world as a van
passed. The energy from it was unforgettable. We remarked to each other that we
would not like to be on top of the hill when he gets there. Next morning Radio
Devon News mentioned that a person had committed suicide on Kit Hill and was
asking if anyone had any information to contact Lostwithial Police. I did give
full particulars of what we had seen. The response was, ``Goodness! You are
spot on. How did you report that information in such a short time?''. I told
them that the energy coming from that van was disturbing. I could not help but
see there was big trouble of some sort.

Whilst talking to the Police, let's talk of the policing of Thorn Moor, Cheriton
Bishop. The area policeman was a typical comic-book build, tubby, red-faced,
and kindly man named PC Tonkins, who lived in the Police house at Tedburn St
Mary. But to me, a small boy was always in the right place of minor offences
committed by small boys. One day my friend did not have a bike so, not wishing
him to walk, I gave him a lift. We arrived with David sitting on my rear
carrier. The shock of seeing PC Tonkins cycling towards me caused me to lose
control and run up a very steep bank and fall backwards with the bike on top of
both of us. PC Tonkins stood in the road and exclaimed that that demonstrated
the reason for not riding two people on a bike. He waited to see if we were OK
before continuing his journey. Luckily it was outside of David's house and I
continued to cycle home. Sometimes we would see PC Tonkins ride past our gate
on his way to his favourite cider cellar, and later in the day riding in a car
with his bike on the back having a very necessary lift home. Some years later
PC Tonkins retired and was replaced by a PC Cannon, immediately renamed by the
locals as ``bang bang''. Big enquiries were made about his character. I wanted
to know whether or not he let minor deeds go unnoticed. I also wanted to find
out whether he enjoyed paperwork, as if he did this would not be good --- he
would probably make out a ticket for blowing our nose. One evening, when I was
aged about 16, I was walking quietly beside a hedge leading to a gate to the
road. As I got to the gate I checked again that the gun I was carrying was
empty and left open. Prior to opening the gate and stepping on to the road I
was confronted by PC Cannon, or should I say Bang Bang. He had remained hidden
until I stepped on to the road. He greeted me with the following: ``Let's see,
your name is Michael?''. ``Yes sir'', I replied. ``Let's see'' he said ``Is
that a 12 bore shot gun?''. Again I replied, ``Yes sir. It needs a shot gun
licence like this one'', pulling the very item from my top pocket and handing
it over. PC Cannon said. ``Ah yes. I thought you of all people would have
one.'' I thought, I expect he was hoping that I did not. I watched him read
through the form and I thought that he was a paper man so nothing out of place
will go unnoticed so watch out farmers in this area. A rumour circulated that
this PC had been relieved after the war as a prisoner of the Japanese Death
Railway. I thought that if that was the case, I would forgive him his general
unforgiving conduct.

Some years later my friend George Gillard and I were asked by PC Alan Parsley of
Whiddon Down if we would become Special Constables to support him with traffic,
or, as happened at the time, check points following convicts' escape from
Dartmoor Prison. This we both did and spent many winter nights having training
at Moretonhampstead Town Hall. One evening when a member arrived late he told
us that President Kennedy had been assassinated. We all stood in disbelief. I
was called to help a regular Constable with a road check at Beter Cross on
Dartmoor commencing at 7pm until midnight. We sat in my car as the night grew
cold with heavy fog due, so we could hear a car from miles away and would get
out of the car standing on each side of the road and wave our torches at the
car to stop. Some would not wish to stop as our torch lights would reveal their
passenger to be someone other than his wife. Not our care, so we thanked them
for stopping and waved them on their way. In the late 1950s I was led to
understand convicts would escape to pay back a debt to the tobacco baron, and
the longer they could stay out and cause disruption the more the debt was
reduced.
