%!TEX root = ../thorn-moor.tex


\emph{Weather and action required by farmers in the month of March.} When the
blackthorn is covered with white flowers it is called a blackthorn winter
(period of very cold east winds requiring overcoats). East wind is called a
lazy wind, too lazy to go around but blows straight through. When the
blackthorn blossom is white, till the barley day or night. March winds and
April showers bring forth May flowers.

\bigskip
\noindent Advice to youth in respect of clothing: never cast a clout until May
is out.

\bigskip
\noindent
\emph{Regarding Dartmoor weather.} When I was a young man living at home before
marriage --- location Thorn Moor Farm, in the Big Pennypark field --- looking
across the Devon countryside at 15+ miles directly into Causdon Beacon, if the
moor looked very clear and close whilst sitting at the bottom of Big
Pennypark, rain would develop in the next few hours or days and the heavy laden
rain filled-humidity would cause magnification of the moor. If the moor looked
far away and purple we would receive fine dry weather in a period of time. Home
wind blowing east: dry day; wind from the west: warmer and wet.

\bigskip
\noindent When Kirton (Crediton) was a busy market town, Exeter was nothing but
a fuzzy down.

\bigskip
\noindent Instructions from a farmer to his workmen: ``Boy! Rin, boy ride, boy
rin and tell the doctor Mrs been and poked the cow's horn in her eye!''

\bigskip
\noindent Two pigs in a house by themselves do better than one together.

\bigskip
\noindent Farmer looking over his crop of potatoes in the rain: ``this shower of
rain will spoil the little potatoes''.

\bigskip
\noindent Trouble with clocks going forward for summertime: it gets late early!
