%!TEX root = ../thorn-moor.tex

Some 18 months later I was promoted to the position of Master Mechanic (I think
a position that is held in the USA Mechanical Engineering world), and within a
few weeks of supervision at the engine rebuild team we had a classroom in which
we rebuilt the turbochargers and fuel pump injection units. Every engine built
up to 977 and D6 size was power tested at Exeter Depot test bed. Larger engines
such as the D8 and D9 were shipped to Cannock Depot for testing. You can
imagine the nervous anticipation when unwrapping the engine on its return,
looking for oil leaks or any other fault normally marked in very bright red pen
or felt tip --- nothing was spared. (It was called Inter Depot Competition, or
`one upmanship'). At the bottom far location was a very special bay housing for
heavy duty track press, necessary for the rebuilding of tracks and pin-and-bush
turning. Three bays were used for whole machine rebuilds. The welding team
occupied a bay for the use of rebuilding various bucket track frames at the
rear of the workshop, which also held a machine shop and track roller
reconditioning unit. Upstairs was a works canteen that management shared. All
meals were cooked on site by the resident cook named Olive, who was aged 45ish.
The rest of this floor contained the Depot Manager's office, and the General
Office containing the Service Manager and Sales Team. For some years the Depot
operated a rebuild system called `Bondedbuy', where a low loader lorry with
Caterpillar D8 or D9 would arrive from an open-cast coal mine that could be
described as end of life. The unit would be power washed in the yard, pushed in
to the workshop, and completely dismantled, with each unit overhead craned into
their respective bays for rebuilding. The rebuild was of a high standard as
they were marketed with a six-month warranty --- the same value as marketed by
the manufacturer when sold new. Two-week visits to the Caterpillar factory
CAT School certainly enhanced the practical knowledge gained in the last two
years.

One Friday afternoon I was called to the Field Service Manager's office. He said
that, as I had been in the Field Service world of agricultural engineering,
would I consider taking the position of Field Service Engineer, as experience
was invaluable with the proposed construction of the M5 from Bristol down, so
would I give the matter careful consideration with my wife. Careful
consideration being the key element to a Field Service Engineer position with
Caterpillar --- that was unusual. Holiday dates would remain as requested other
than service problems. It was hoped that I would respond any time of the day or
night, Bank holidays and weekends, as the emergencies of the of job required.
We had to take into consideration that at this time Caterpillar had probably
the most experienced mechanics, together with components, labour, and also
living near the motorway. The Salesman sells the first machine. The Service
sells the rest! Monday morning found me starting the changeover from workshop
to Field Service man, an event that did not start off in a manner I would wish:
all tools and equipment having been collected together and transferred to the
field service van. I was dropped off to collect my Ford Transit van from the
paint shop which was situated near Polsloe Bridge in Exeter, unfortunately. I
say unfortunately because I misjudged a right-hand square brick corner and
touched the side door. I did not stop thinking of the old saying ``Do not stop
to look or you will never start and go again''. When I arrived back at the
Depot I lined my van up with the other two. The scene of the people changed
into a circus. The Service Controller continually kept looking into the filing
cabinet. Some people smirked and slid away from the scene. When I walked around
the van and examined the damage it looked like it had been in a Morris dance
with a chainsaw. It was then I heard a cutting voice say ``You will have to
manage for a week or so'' and manage I did, feeling very confident with the
repair required. All components and gaskets are manufactured to a high quality
and give a higher confidence in repairs.

The first taste of the nature of the work to be experienced in this area was
Cullumpton By-Pass. We serviced and repaired a spread of D8 and 977 machines
used in the burrough pit. Some CAT 631 scrapers were added to spread the
landfill. Running repairs were saved until Saturday mid-day, when the site had
afternoons off. We would be joined by two or three men from our workshop, and
repairs would commence from 12pm, working through the night until ready for
the commencement to work Sunday morning. If extra flood lights were required,
we would burn the gas off the welding set, keeping the nozzle free of carbon
build up. One day we were informed that a delivery of new Russian earth-moving
equipment was due and would outperform. In due time a delivery of bulldozers
quite obviously came out of the factory which had a small quantity of military
design, then had a bulldozer system, several off-road lorries and other
equipment that had started life as military. These units started and revved up
to 6000 rpm and smoked off unburnt fuel and oil until the operator disappeared
from view and ran round the site at speed with questionable stability as the
track idle was not designed for civilian work. The shattering noise of site
machinery gradually diminished and returned to the steady pulse of the original
equipment as it became apparent that the new Russian machine was lying around
waiting for a part replacement. Time marched on together with the construction
process of the by-pass.

Over the period of time I followed up the marine engine problems as a natural
progression of my engine bay rebuilding days. This was enhanced by an engine
power output and turbocharger week course at the Caterpillar factory located in
Glasgow. My first impression as a country boy was riding on the top deck of a
bus and, whilst riding through an area called Galowgate, seeing the doors of
each pub open wide and staff washing the stone floor free from stale spilt beer
out into the gutter --- the acrid smell of stale beer running through the
length of the bus is a constant reminder of another world. A visit I made to an
oil-pumping rig in Dorset presented real history to me. The Caterpillar engine
that was the power unit to this oil-pumping rig was a very old unit --- a
Caterpillar 2000. This was approximately the size of (being the forerunner of)
the D8 power unit with external push rods. The fault was a severely leaking
water pump, constructed with one main through shaft of 5/8'' in diameter using
a bronze calibrator adjusted to tighten a hemp-impregnated graphite rope gland
onto the shaft: as the unit had been running a lifetime with, I suspect,
regular service adjustment, the main shaft diameter of 5/8'' had worn down to
1/2'', eliminating any further adjustment to be taken. So a rebuild of parts
was available and they were from stock in the USA. Parts were delivered within
a few days and the pump rebuilt. A bigger surprise greeted me on my return to
the oil rig. Records show that the engine had been in full use after D Day. It
had been used to power the system known as PLUTO that pumped the oil and fuel
at the bottom of the sea bed across the channel to supply all our allied forces
and had been removed after hostilities to continue in peacetime work. What a
wonderful history to find by chance.

My position as a Field Service Engineer encouraged the company to send me to
Geneva, the then CAT School of Europe, located at the foot of Mont Blanc. The
tutors were Swiss and the class consisted of dealership engineers from all over
Europe --- a Spaniard, two Belgians, a Danish, a Phillipino, and myself.
The first week was power shift and torque converter transmission; the second
week was loader hydraulic system; but the most interesting course for me was
the two weeks on boat hull speeds propeller-to-power calculation. To touch on
the above, there is no doubt it is an exact science in itself but a
miscalculation on each would bring heartbreak! I must confirm that the Courses
at the Geneva CAT School were booked for many service dealerships' service
engineers. It is said that a little knowledge in the wrong hands can cause
problems. This saying came to mind when I was called to Southampton to carry
out sea trials on a 55' MV vessel using twin Caterpillar D434 engine. The boat
itself was on a new build completion up the Hamble in a boatyard run by Robin
Knox-Johnson. After the general inspection and test the boat was taken into the
Solent for trials and was near to completion when it was found that full power
was not achieved. We were down 50 rpm. All tests to the turbo charger indicated
that engine was up to performance. I said that the propellers pitch or diameter
should be checked for correct size. This comment of mine stimulated the
question, ``could you explain further?''. My (unqualified) response was that it
could be that the diameter of the propeller needed reducing by 1/2''. So I
returned to Exeter in the full knowledge that the propeller would be
reinvestigated and I would be recalled on completion. Eventually I returned to
the Hamble for what should have been a final sea trial. The result of the full
power --- yes you've got it --- had lost a further 50 rpm and was down by a
total of 100 rpm. Another lesson in life --- a little knowledge in the wrong
hands can cause problems! The Propeller Company had lost the services of the
propeller calculator and had reverted to looking up past history of boat sizes
in relation to propeller size, and several factors would not have been taken
into account. So this boat had been installed with too small a diameter
propeller which caused cavitation that introduced extra power required of the
engine, hence the loss of 50 rpm in the fast speed run. Removing a further
1/2'' magnified the power required. It was the marine course in Geneva that
revealed all in later months.

This course was to furnish us with details of the correct question we should
ask, such as ``what weight did the boat finish up with at the build finish'',
as well as teaching us how to use the slide rules supplied to calculate prop
size. Regarding this we could stand our ground and respond with the useful
information required. The first day of the course was spent analysing engine
output or lack of it, the remainder explaining that a hull designed to a
particular shape and long ton weight could not exceed a certain speed --- for
instance, a hull with a power of 500 hp unit with a maximum design speed of 7
knots will never go over that speed if the engine is replaced with a 750 hp
unit. Also correct propeller size is vital for this. We were presented with a
special set of slide rules and taught (with difficulty) what vital
specification to ask for before coming to work out propeller size on the slide
rules. ``Why go to such depths?'', you might ask. Well, if you are the CAT
engine test engineer on the sea trial and the engine reaches full rpm and
power, everyone gets a pat on the back and completes the sea trials with a big
smile. If not, and the rpm was below specification, open warfare would erupt in
the wheel house between the designer, boat builder and the CAT man --- yours
truly. The weekend came and no let up. We were paired up, I with the Spaniard,
and given a problem that needed a propeller calculation using the slide
rules --– heaven forbid that I was always lagging behind the very switched-on
Belgian and needed a night's rest to think things over. We got to the hotel and
my partner, the Spaniard said ``Bye Mike, I am here to enjoy myself --- see you
Monday morning''. Saturday morning with the warm sun shining on me I walked,
with my homework, the 200 meters from the hotel to the Jet d'Eau on the lake as
I sat beside this lovely lake environment looking at the jet, and I was
reminded of the jetting on the lake surface to the top of the nozzle, now 140
meters with 7 tons of water. Working alone on the problem I had a wonderful
time as a local old man offered me a ride in his boat around the lake. It ended
up a very delightful 2 days (but you will soon feel sorry for me). Back in
school, questions were asked to the teams of two, and it arrived at us before
any questions could be presented by the tutor. My Spanish friend said loudly,
``As Mike is an English officer and a gentleman, I left him with the problem.''
Well, I took the full force of the questions and corrections had to be made. I
felt I came out wiser but somewhat sand blasted. Out of this became a better
understanding and the desire to think about imagining faults and possible
remedies when I had spare time to dream.
