%!TEX root = ../thorn-moor.tex

William James Saffin, my Great Grandfather, lived at Poleford between
Crockernwell and Cheriton Bishop. He was a stone cracker, as in his early days
the A30 road was not covered in tarmacadam (it was probably not invented then)
but was a hard earth surface with the covering of duck egg sized stones that
would gradually break down to small granulated rubber under the iron rim of
various stage coaches, farm wagons, and carts. In fact, Great Grandfather
remembered a special occasion when a man on horseback rode through with one of
the first motor cars, which drove through with a red flag bearing the words of
a warning of mortal danger and ``beware'' hanging on its side in the form of a
notice. This was the first motor car to drive through the A30 road. Dad said
there was one story Great Granddad told in solemn mood, which was the appalling
attitude of the aristocracy towards the working man experienced by him. The
story goes that Great Grandfather had finished work for the day and started to
walk home. After a time of walking he observed a gentleman smartly dressed and
riding a horse. As he drew nearer my Great Grandfather gave a happy smile. His
greeting was received with a very cold, stony expression and he was taken
completely by surprise when this individual gave him a full and hefty lash with
his horsewhip across his back with a force draining all his strength, causing
him to fall on his hands and knees. Fortunately he had been an active working
man and was in fairly good rugged bodily condition. His recovery was quite
quick and as he started to rise to his feet he selected a good size ragged
stone. The horseman had lingered to watch my Great Grandfather's plight while
he was lying among the dust and stones, and turned away to continue his
journey. My Great Grandfather threw his stone as hard as he could, hitting the
back of the horseman's head. He fell off his horse with one foot tangled in the
stirrup. He regained balance, remounted and kicked the horse to continue his
journey. There was never a word spoken about the savage incident between my
great grandfather and the gentleman horse rider. I have often thought of the
unsavoury incident and wondered what blackness must have been in the heart of
the rider --- the only two men in a lonely country road, one giving a smile to
then receive a whip lash with no other people present to witness the violence.

Going back to my grandfather John Preston Butt, known to all as Jack Butt, he
had talents which were greatly admired by me as a small boy. Naturally my
awareness expanded as the years went by and I immediately acknowledged his
skill as a thatcher and his ability in making roof spars, straw ropes, and bee
skips. Then one morning I was aware that a letter addressed to Jack Butt had
arrived by postman Mr Hawkins (Smiler). Mother read the contents to Grandfather
whilst he sat on the Devon settle listening with great concentration --- we
must remember that he had limited writing skills and even less reading
abilities. I recall Grandfather saying ``I was taught how to read and write by
a man who would say if I stopped (because I did, not knowing the word)
\emph{read on boy} thinking thicky man's daid'' (dead). Irrespective of this
limited tuition he was able to write on a slate with a long carpenter's pencil
the customers' names and numbers of spars made and bundled for them, and total
up the pounds shillings and pence. Following on from the letter reading I
found Jack Butt dressed in his second best waistcoat and carrying his bowler
hat and three or four newly cut hazel sticks waiting for a taxi. Naturally
after his departure I made enquiries to be told that he had gone to do some
water divining for someone who wished to dig a new well. It was some years
later that he showed me the technique telling me that only I would know when I
was ready. I have never committed myself to divine water for anyone as in my
opinion it could cost money. For my own peace of mind I have divined and found
a healthy flow of water running under Cowick Street shopping centre car park
in Exeter --- maybe a big drainage. I was also asked if I would confirm the
finding of two professional diviners who had reported the finding of a large
spring flow of water across the road on a driveway between Newton Abbot and
Penn Inn. I attended and to my horror and disappointment could not get any
reaction from my hazel sticks but walking 50 yards away a great reaction
against a stone wall. I reported my finding and disappointment to have lost
the little gift I had! So I got further help. Approximately a month later I
received a phone call from the land owner saying that I should not be
disappointed as they had carried out drilling probes in the drive which
confirmed that I was correct that there was no water under the driveway but
they had found water 50 yards away, against the wall which was once used as a
communal water tap from a well, so I felt very pleased. In general, a letter
would arrive giving a communication for a date at Grandfather's workshop for a
taxi car to arrive, and the visitor going into the shed with the door tightly
closed for privacy time. After some years it was noted that a few small
parcels addressed to Jack Butt and tied up with string with the knots sealed
with sealing wax would be delivered and opened to reveal an ounce of Digger
shag tobacco or may be two. Grandfather would never give away anything about
charming warts. I was then told that it must be passed from male to female so
Mother received the information from Grandfather but did not feel happy to use
it so passed it to me. I charmed many warts over the years but did not sell
the gift, although I have received great pleasure when thanked after the warts
disappeared. Folks might laugh and doubt, but if they get a wart on their nose
they might try, if it won't go, to see me.

One more historical story: the blizzard of 1963. The blizzard started just after
12 o'clock on the morning of Boxing Day. I had milked the cows and all was in
order, as I thought it was just another flurry of snow. Mother and father were
stocked up with food and normal warmth. Dad was recovering from an operation at
this time. I then drove to Brampford Speke to visit Vera for lunch. As I was
visiting the love of my life, lunch went into tea time, and at 9.30pm with no
regard to the weather I looked out of the window and dressed for the journey
home. You can imagine the horror and deep feeling of irresponsibility when I
saw the extreme blizzard conditions knowing that my mother was at home with my
unwell father in a very isolated house and me 15 miles away. At that time I was
driving a Ford 1 Courtina with a three speed column gear change and small
wheels which were most unsuitable for traction! However, I started to drive
with a passion knowing that I must get home to my parents. So I drove
zig-zagging through snow drifts while keeping the speed up to power through the
drifts. People who have driven through snow storms have experienced the feeling
of uncertainty when driving uphill, along the flat, or downhill. I experienced
this ungodly sensation. When I reached the top of Burridge Hill the engine
gained speed and I identified that I had gained the ascent and was descending
far too fast to negotiate the T junction 100 yards ahead, so I did my best to
reduce speed, well knowing that I was going to contact the hedge, which I did.
Fortunately only the front wheel ran up the bank leaving the car sitting with
the front wheel sitting in the hedge. So, knowing the rear wheels would not
grip, I pulled my raincoat out of the boot of the car and placed it under the
rear wheels, jumped back into the car, shut the engine down, put the gear into
reverse and pressed the start button and the car wound its way back off --- a
very handy trick that modern cars are unable to do. I drove off leaving my
raincoat in the road as the weather was too harsh to linger. I drove
continually zig-zagging through small snowdrifts knowing that after Tedburn St
Mary conditions would become much worse. Fortunately my judgement could not
have been further from the truth since, in reality, as the A30 became more
exposed to the strong winds tearing out of Dartmoor, the snow was being blown
over the hedges, leaving me a reasonably clear run until I turned off onto the
Yeoford road. The reality of the blizzard turned my progress belly up. The wind
had blown the snow from the open area of Red Ridges over the hedge to cause a
massive 11' high snow drift across the road into the entrance to Glebelands
Estate. Just down the nap from the Village Hall everything was a white out and
this obstacle could not be seen, so I went into it at a fairly steady speed and
I think the car climbed into the drift at about 8' high, first a rapid
deceleration, then to a complete stop, and also in complete darkness as the
headlights were buried in the mass of snow. The door could not be opened as we
were wedged in the drift so I wound down the window and climbed out on to the
snow. The next big shock was when I sunk deeper down in the snow and, with some
degree of panic, I detached myself from the snowdrift leaving my car to rest
with windows open and with only the roof and windows visible (but to be covered
later with the falling snow). I switched on my trusty Pifco lamp and set out to
walk home. After I had walked, sometimes through knee deep snow, I became aware
of a clanking noise around ears, so, feeling my head, I was then aware that the
snowflakes landing on my warm head had melted and, whilst running down my hair,
had frozen in the cold harsh wind. How unwise not to be wearing a hat. On
arriving home I was relieved to hear both my parents' calls who responded with
reassurances that I was home. Next morning found the snow level with the
windowsills, and it was still falling heavily. It was obvious to me that Thorn
Moor, Little Moor Farm, and other isolated farms were cut off. Whilst feeding
the cattle I reflected on my very lucky and dodgy drive back the previous
night. I thought what a stupid selfish bull-headed individual I was, well used
to knowing the weather in store with a responsibility for a sick father and a
very anxious mother and with animal feeding and milking to be done. However I
settled down to the running of Thorn Moor under blizzard conditions.
Fortunately it was a holiday break so work was out of the question. I cannot
remember having a day when it didn't snow but it eventually stopped. I then
walked to Checkers shop at Cheriton Bishop for provisions, and rather than
return on the strength-sapping powdery snow I decided to walk to Crockernwell
via the A30 and Hooperton Cross, which was much longer and involved an extra
walk of about two miles in blizzard conditions through knee-high snow. The
worst was trying to breathe in the corridor of very fine particles of
broken-down cold snow that had filtered through. I set up my very unsuitable
car for snow use, adjusted the rear tyres to half normal pressure and front
tyres a little higher than standard. I wrapped a rope around the rear drive
wheels to connect blocks in the boot. I gave the farmer living at Little Thorn
Farm £1. He then towed me through the heavy snow to the stone landing. My Ford
was left unlocked because of the extra low temperature --- a lock would never
open with the boot facing Dartmoor's extreme weather. To keep the engine clear
of snow blown by the wind each morning I would walk from Thorn Moor then drive
to work, so never was a days' work lost, as we had plenty of water pump air
vessels to cast weld together with a couple of Allis Chalmers engine cast
blocks to also weld. After some days I was aware that the road had hard packs
of snow so the car was driven home. After some more days it was noticed that I
could see over the hedges and walls, so the thickness of the packed ice was
approximately one inch deep and I was very aware that the thaw would cause
problems when the surface broke up. So for some days over the thaw the car was
left at Hooperton Cross and I walked to and from it each day until the normal
surface returned.
