%!TEX root = ../thorn-moor.tex

The Propeller Company had lost the services of the propeller calculator and had
reverted to looking up past history of boat sizes in relation to propeller size
and several factors would not have been taken into account. So this boat had
been installed with too small a diameter propeller which caused cavitation that
introduced extra power required of the engine hence the loss of 50 rpm in the
fast speed run. By removing a further ½'' magnified the power required. It
was the marine course in Geneva that revealed all in later months. Another
occasion I had worked continually without a day's break and experienced a
desire to be freed up for the weekend. So you can imagine the anticipated
build up of the break was joyous and by 5 o'clock on that Friday afternoon I
jumped into the van and was about to insert the ignition key when the field
Service office door burst open and the supervisor Joe burst through with a very
serious red face and said, `` We have a problem --- a new power boat has been
delivered to St Hellier, Jersey this morning and whilst demonstrating its
manoeuvrability to the good and the great in the harbour had suffered a major
marine gear failure (and yes you've got it in one) it is so late that we are
unable to book you a flight out of Exeter so you are booked to fly out of
Southampton first flight in the morning. So it's an early start next morning
not a lie in but up with the blackbirds to Southampton. My brief received
between Jersey Airport and St Hellier was equally intense. Plans were in
position for a birthday party celebrating the arrival of a new Fairy boat. All
Fairy boat owners had been invited to assemble with their boats to the inner
harbour of St Hellier on Saturday morning with mooring drinks at the Yacht Club
and when refreshed to set off in a group to St Malo in Brittany for the weekend
celebration. In fact at this moment in time the party was assembling and was
advised that there had been a slight hitch and that the Cat man would be
arriving from England in the morning and should correct the problem. Then all
will be go! Using todays term no pressure then. When I got to the waterside
and observed the collection of power boats moored alongside in anticipation of
the forthcoming jolly together with the owners and ladies one realised how far
the original event had travelled before reality had set in. Particularly when
the South Pier manager, Gerald, walked over looking ashen and said, ``I hope to
God you can pull this out of the bag''. We shook hands and I went below to
examine he bowels of the problem. It was quickly apparent that this was no
quick fix as a clutch plate had shattered and centrafugaly spread and jammed
the drive astern gear solution, remove marine gear and needed to be returned to
Exeter Depot for rebuild. The news was received with much disappointment and
fury all directed at the manager. Fortunately it was decided to continue with
the party as all were gathered and full arrangements had been made. I continued
to remove the marine gear for transportation greatly assisted by the South Pier
shipyard which had a well organised team. The weather over that period was
exceptionally hot and sunny. I unwisely decided to work shirtless so by the
afternoon a bottle of calamine lotion was purchased and used to cool my very
hot back. By Monday the gear was despatched via a boat motoring to harbour. Due
to local fog the airport was shut down so I was taken to the harbour and
located a coaster that was on a passage to Southampton that night. So I signed
on as a supernumery with a Captain Edwards as a non-working crew member of MV
The Loon Fisher. I spent part of the night sleeping in the sick bay and part
admiring the fortitude of the Bellman standing out in the bows dressed in storm
clothing and covered in damp fog particularly constantly peering into the mist
of impending danger. We docked at Southampton and I must have looked odd
walking alone between the shipping containers carrying a very heavy holdall
looking for an exit. I was quickly arrested by a Customs officer and taken for
questioning and a strip search. A degree of satisfaction spread over me when I
was standing in pants only whilst the Customs officer passed around the bottle
of calamine lotion for tasting and judgement proved satisfactory and I returned
with the marine gear that was rebuilt and taken back to St Hellier and refitted
without further dramas. Life continued with repairs and warranty work as
required with normal six weekly marine work continuing in Jersey. I was then
called to another fast patrol boat, a 45' Keith Nelson hull which was used for
river police and a pilot boat. I was engaged on a top overhaul to the engine,
removing cleaning, replacing and sealing air intake inter coolers. One
turbocharger needed rebuilding so an exchange unit was flown over. This tale
is not about engineering or technology. It is about things that unexpectedly
happen that cause disruption, I think you must stop and take time to think how
do I get this back to the norm with minimum action. Taking up the story the
replacement turbocharger was flown out and delivered to South Pier shipyard.
Douglas Park, a skipper to the Duchess of Normandy, an area patrol vessel,
agreed to deliver the unit to me by using his harbour runabout. To recap the
scene, all boats in the harbour were normally afloat and went aground at low
tide of 2-3 hours. To set the scene it was a beautiful afternoon with a spring
tide washing up to the top of the tide at approximately 7 o'clock with sun
shining across the water surface. Douglas threw a rope across to me with the
words, ``Lift on that one when ready as I have tied the package in a tight rope
parcel until the package was midway between the boats and open water. Suddenly
the rope went slack followed by a glug with a multitude of bubbles rising to
the surface. A stunned silence between us followed by me asking the question,
well knowing the answer, but it gave me time to recover. The response was 2 am
tomorrow morning would see the tide recede down enough for me to walk out in
the harbour and retrieve the package. So that evening I had a very uneasy
dinner at Mrs Cornish's and did not change out of working kit but walked across
to the Folly and had a couple of drinks with the boys and then went back to my
room and watched the clock. At 1 am I put on a sturdy evening coat and walked
out into the night and made my uncertain way down into the dark harbour. You
could imagine every shadow that bounced off the dark pillars in the moonlight
was a demon ready to cause me harm. So I quietly made my way to a pillar and
stood very quietly in the shadows keeping my eye on the surface of the harbour
floor and quickly walked out and retrieved the package putting it in my sack
and walked back to the Guest House. I have wondered how I would have responded
to the questions which might have been asked by a patrolling policeman. A lone
man walking through the streets of St Hellier carrying in both hands a valuable
turbocharger from power boats in the middle of the night I reason I could have
spent the night in a cell until the truth was confirmed. However next day the
unit was dismantled, washed of seawater and relubricated and installed to
continue            to work as required. Between my marine work I
was attending the motorway section of the M5 and M4 carrying out maintenance as
required working in that location a week at a time from Banwell to Edith Mead.
If parts were required we collected it in the early morning from our Melksham
Depot. It was noted that in the warmer summer week the flies and mosquitoes
multiplied from the wet reed beds along the flat marshy land caused much
stomach problems among the machine operators working in that area. At that
time I was called to fly out to Malta to correct lack of boost pressure in a
turbocharger with intercooler in use in the boat which had limped down the
Adriatic to arrive in Silema Harbour. It was an 80' yacht with a beautiful
name, The Phyllis Serena. I went aboard this wonderful pristine boat into the
bowels and opened a round hatch into the black hole of the engine to be greeted
by a noise like jingle bells. I found myself standing knee deep in a mixture
of empty one pint size oil tins and pull off slips. It became apparent that no
maintenance or kindness was ever given to this engine. However the unit
required an oil top up. Someone would find a car garage and purchase an armful
of one pint oil (any oil) and return to the engine and fill as required and
throw the empty can together with tin foil slip on to the engine room floor and
walk away. Sadly this state of affairs had caused the internal components to
work in mud so the turbocharger together with the intercooler was removed and
taken to a workshop and I dismantled and rebuilt after a thorough degumming.
The engine sump was removed and the mud like black oil scraped out and the
engine rebuilt. To think it cost a lot of money to send me from Exeter to
Malta to correct this problem and someone had been employed as an engineer to
this boat yet maintenance never existed over many years. Let me tell you about
the workshop that had kindly been offered for me to use together with all the
tools needed. I was driven then by the boson, Charlie Greengrass in a Mini
Moke. My first impression was of quite a large galvanised corrugated shed with
the roof completely covered in live and growing course green grass. When I
opened the door and walked in I saw the green grass had caused the roof to rust
through. When it rained the rain would run through the funnel shaped hole and
hit the centre of the workshop floor. Modification had been made to the
system. A hammer drill had been used to cut a gutter from the leaking water
across the floor under the door and across the outside pavement and in to the
street. The workshop was empty and clear but around the wall hung new cabinets
containing thousands and thousands of pounds worth of USA snap-on tools not
used and brand new.
